---
title: "Salmon Species Caught Analysis"
author: "Adam"
date: "`r format(Sys.time(), '%d %B, %y')`"
output: html_document
---

---

This analysis is of the number of salmon caught over time. Whilst both of these plot's code are relatively simple, with the initial plot it is a simple line plot of capture against the year, coloured by the species. In the final plot, I opted for an area chart rather than a line chart, and also changed the theme, and the position of the legend.

---

```{r setup, include = FALSE}
knitr::opts_chunk$set(fig.align = 'center', out.width='80%', echo = TRUE)
```

```{r imports, include = FALSE}
library(tidyverse)

load(file = 'fish.RData')
```

## Final Plot
```{r finalplot, echo = FALSE}
ggplot(fish.tidy, aes(Year, Capture, fill = Species)) +
  geom_area(aes(group = Species), alpha=0.85) +
  labs(title = 'Amount of Species of Salmon Captured Over Time',
       y = 'Number Captured') +
  theme_classic() +
  theme(legend.key.size = unit(0.5, 'cm'),legend.position = 'bottom')
```

---

## Initial Plot
```{r initialplot, echo = FALSE}
ggplot(fish.tidy, aes(Year, Capture, color = Species)) +
  geom_line(aes(group = Species))
```

---

## Code

### *Code for the final plot:*
```{r finalplotcode, eval = FALSE}
ggplot(fish.tidy, aes(Year, Capture, fill = Species)) +
  geom_area(aes(group = Species), alpha=0.85) +
  labs(title = 'Amount of Species of Salmon Captured Over Time',
       y = 'Number Captured') +
  theme_classic() +
  theme(legend.key.size = unit(0.5, 'cm'),legend.position = 'bottom')
```

### *Code for the initial plot:*
```{r initialplotcode, eval = FALSE}
ggplot(fish.tidy, aes(Year, Capture, color = Species)) +
  geom_line(aes(group = Species))
```

