---
title: "Investment Report"
author: "Adam"
date: "`r format(Sys.time(), '%d %B, %y')`"
output: 
  html_document:
    toc: true
    toc_float:
      collapsed: false
params:
  country: Brazil
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(fig.align = 'center', out.width='80%', echo = TRUE)
```

```{r data, include = FALSE}
library(readr)
library(tidyverse)
library(knitr)

investment_annual_summary <- read_csv('investment_annual_summary.csv')
investment_region_summary <- read_csv('investment_region_summary.csv')
investment_services_projects <- read_csv('investment_services_projects.csv')
```

## Datasets

### Investment Annual Summary

The `investment_annual_summary` dataset provides a summary of the dollars in millions provided to each of the following regions for each fiscal year, from 2012 to 2018:

- East Asia and the Pacific
- Europe and Central Asia
- Latin America and the Caribbean
- Middle East and North Africa
- South Asia
- Sub-Saharan Africa

```{r investment-annual-summary}
ggplot(investment_annual_summary, aes(fiscal_year, dollars_in_millions, color=region)) +
  geom_line() +
  labs(title = "Investment Annual Summary",
       x = "Fiscal Year",
       y = "Dollars in Millions") +
  theme(legend.key.size = unit(0.5, 'cm'),legend.position = 'bottom')
```

```{r tables, echo = FALSE}
kable(investment_region_summary, col.names = c('Region', 'Dollars in Millions'), align = 'cc', caption = 'Table 1.1 The The total investment summary for each region for the 2012 to 2018 fiscal years.')
```

### Investment Projects in `r params$country`

The `investment_services_projects` dataset provides information about each investment project from 2012 to 2018. Information listed includes the project name, company name, sector, project status, and investment amounts.

```{r country-investment-projects, warning = FALSE}
country_investment_projects <- investment_services_projects %>%
  filter(country == params$country)

ggplot(country_investment_projects, aes(date_disclosed, total_investment, color=status)) +
  geom_point() +
  labs(title = "Investment Services Projects",
       x = "Date Disclosed",
       y = "Total IFC Investment in Dollars in Millions") +
  theme(legend.key.size = unit(0.5, 'cm'),legend.position = 'bottom')
```


### Investment Projects in `r params$country` in 2018

```{r country-investment-projects-2018, warning = FALSE}
country_investment_projects_2018 <- investment_services_projects %>%
  filter(country == params$country,
         date_disclosed >= "2017-07-01",
         date_disclosed <= "2018-06-30") 

country_investment_projects_2018_total <- country_investment_projects_2018 %>%
  summarize(sum_total_investment = sum(total_investment, na.rm = TRUE)) 

ggplot(country_investment_projects_2018, aes(date_disclosed, total_investment, color=status)) +
  geom_col(size=3.5, alpha=0.5) + 
  labs(title = "Investment Services Projects",
       x = "Date Disclosed",
       y = "Total IFC Investment in Dollars in Millions") +
  theme(legend.key.size = unit(0.5, 'cm'),legend.position = 'bottom')
```

The total investment amount for all projects in `r params$country` in the 2018 fiscal year was `r country_investment_projects_2018_total` million dollars.
